---
title: "NaturalDisasterPortfolio"
author: "Jacob Organista"
format: html
---

```{r}
library(tidyverse)
library(readr)
```

```{r}
X1900_2021_DISASTERS_xlsx_emdat_data <- read_csv("~/Desktop/STAT 210/Final Portfolio/DISASTERS/1900_2021_DISASTERS.xlsx - emdat data.csv")
```


*Data Descriptions:*

1. This is a recorded Worldwide Disaster Data Set spanning from 1900 - 2021.
2. The data source that I will be using for this portfolio is a table of worldwide disasters that include data about where and when the event was, what type of disaster was it, and how it affected the area which includes death tolls, injured numbers, damage costs, etc. 
3. The research questions that I am interested in determining are:
  a. Does the country of origin affect the amount of fatalities/ damage costs?
  b. How does the type of disaster differ in the result of the area that it affected?
  
*Data Visualization:*
1. For the final visualizations, I think a bar chart would work best for the first research question with the x-axis displaying the country of origin, and the y-axis displaying the damage costs. For the second research question, I have a couple of ideas which include another bar chart except with 3 portions. The x-axis would show the type of disaster, the y-axis would be the magnitude/ impact of the disaster which could be injured, deaths, damages, etc, and the fill could be the regions. Another idea for this question would be a scatter plot using the same axis or possibly a heat map would be interesting to see with this idea as the rows could be the disasters, the columns could be the region, and the color intensity would represent the impact level (eg. dark red = more deaths). 
2. I think that the charts will highlight and target which disasters cause the worst damage because this gives insight on how to prepare for the disasters, especially the ones that are within the person looking at my chart's region.
3. I don't think that there are any variables missing that are relative to answer my research questions.

*Data Cleaning:*
1. There are many variables within the raw data set that will need to be adjusted such as dates, times, and  ID numbers. There maybe a couple more that I could find along the way but from my initial scan these stuck out to change first.
2. There is a total of 16,000 entries and right off the bat I can determine there is a lot of missing data and NAs for many of the different variables.
3. I think using filters will be useful so that I can look at the variables specific to each of my questions because there is a lot of metadata that isn't as useful such as latitudes and longtitudes when I already have the region and country of origin.
4. I think there are already a lot of variable, even some combining population totals and damage totals so I shouldn't need any new variables.
5. I don't think I need to use any joins as much of my raw holds the information needed in one file. 
6. I think the majority of the variable will be helpful to look at but as the graphics come together, many of the variables will be unused such as when looking at location, I could use the country, region, continent, or city so there are options, but not every one will be used.
7. I think pivoting could be useful to adjust the data for each question so I will be something that I can use to summarize what is needed for each research question.
8. The data comes with many totals but I think that the summary tool will to useful to understand mean occurrences of disasters in their region.
9. Besides the answers to the questions above, I can't think of more ways to adjust the data to better fit an answer to each of the research questions.

